---
title: "Photogrammetric Point Cloud Processing with lidR"
author: "Mateus Tinoco Silva"
date: "17/10/2021"
output: html_document
---



## Introduction
### Lately, several works have been conducted aiming to investigate possibilities of extracting forests’ structural metrics from  Digital Aerial Photogrammetry (DAP) point clouds (PC).  <br />  This markdown demonstrates a simple workflow on lidR for DAP-PC processing and tree metrics deriving.

## Recommended Bibliography

<https://jean-romain.github.io/lidRbook/index.html>  <br /> 
<https://doi.org/10.1007/s40725-019-00087-2>  <br /> 
<https://doi.org/10.1016/j.rse.2018.05.016>  <br /> 
<https://doi.org/10.3390/rs13183655> <br /> 
<https://doi.org/10.3390/rs12213560>  <br /> 

```{r p1, message=FALSE, warning=FALSE, include=FALSE}
r3dDefaults = rgl::r3dDefaults
m = structure(c(0.921, -0.146, 0.362, 0, 0.386, 0.482, -0.787, 0, 
                -0.06, 0.864, 0.5, 0, 0, 0, 0, 1), .Dim = c(4L, 4L))
rgl::setupKnitr()
r3dDefaults$FOV = 50
r3dDefaults$userMatrix = m
r3dDefaults$zoom = 0.75

library(lidR)
library(ggplot2)
library(knitr)
library(rasterVis)
library(dplyr)
library(ggplot2)
library(sf)
options(lidR.progress = F)
knitr::opts_chunk$set(
  comment =  "#>", 
  collapse = TRUE,
  fig.align = "center")
rgdal::set_thin_PROJ6_warnings(TRUE)
options(crayon.enabled = TRUE)
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, which = c("output", "message", "error"))
```

```{r p2, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r p3, message=FALSE, warning=FALSE, include=FALSE}
path_point_cloud <- "data/pointcloud.las"
path_roi <- "data/roi.shp"
plot_crossection <- function(las,
                             p1 = c(min(las@data$X), mean(las@data$Y)),
                             p2 = c(max(las@data$X), mean(las@data$Y)),
                             width = 4, colour_by = NULL)
{
  colour_by <- enquo(colour_by)
  data_clip <- clip_transect(las, p1, p2, width)
  p <- ggplot(data_clip@data, aes(X,Z)) + geom_point(size = 0.5) + coord_equal() + theme_minimal()
  
  if (!is.null(colour_by))
    p <- p + aes(color = !!colour_by) + labs(color = "")
  
  return(p)
}
```

## Loading Data
```{r p4, echo=TRUE, message=FALSE}
roi <- st_read(path_roi, quiet =T)
point_cloud <- readLAS(path_point_cloud) %>% clip_roi(roi)
```

```{r p42, echo=TRUE, message=TRUE}
print(point_cloud)
```

## Point Cloud Visual Inspection
```{r p50, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, include=FALSE, rgl=TRUE}
plot(point_cloud,color="RGB",size = 3, bg = "white",  axis = TRUE)
```


```{r p41, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
plot(point_cloud,color="RGB",size = 3, bg = "white",  axis = TRUE)
```


## Creating a Digital Surface Model (DSM)
```{r p5, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
dsm <- grid_canopy(point_cloud,res=0.5,algorithm = dsmtin())
plot(dsm,col = height.colors(50))
```

## Ground Classification
```{r message=FALSE}
ws <- seq(3, 12, 3)
th <- seq(0.1, 1.5, length.out = length(ws))
point_cloud_class <- classify_ground(point_cloud, algorithm = pmf(ws = ws, th = th))
```

## Noise Classification
```{r p7, message=FALSE}
point_cloud_class <- classify_noise(point_cloud_class, ivf(res = 0.5, n = 50))
```

```{r p8, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
plot(point_cloud_class %>% filter_poi(Classification!=LASNOISE),
     color = "Classification", size = 3, bg = "white", axis = TRUE) 
```

## Creating a Digital Terrain Model (DTM)
```{r p9, echo=TRUE, message=FALSE, warning=FALSE}
dtm_tin <- grid_terrain(filter_ground(point_cloud_class), 0.5, tin(),keep_lowest = T)
```

```{r p10, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
plot_dtm3d(dtm_tin, bg = "white")
```

## Normalizing Point Cloud
```{r  message=FALSE}
point_cloud_norm <- normalize_height(point_cloud_class, knnidw())%>% filter_poi(Z>= 0)
```

```{r p12, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
plot(point_cloud_norm, color = "Z", size = 3, bg = "white", axis = TRUE) 
```

## Analysing a Vertical Transect
```{r p13, echo=TRUE, fig.height=1.5, fig.width=8, message=FALSE, warning=FALSE}
p1 <- c(306775, 8353310)
p2 <- c(306732, 8353250)

las_tr <- clip_transect(point_cloud_norm, p1, p2, width = 4, xz = TRUE)

ggplot(las_tr@data, aes(X,Z, color = Z)) + 
  geom_point(size = 0.5) + 
  coord_equal() + 
  theme_minimal() +
  scale_color_gradientn(colours = height.colors(50))
```

## Create a Canopy Height Model (CHM)
```{r p14, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
chm <- grid_canopy(point_cloud_norm,res=0.5,algorithm = dsmtin(max_edge = 3))
plot(chm,col = height.colors(50))
```

## Individual Tree Detection 
```{r message=FALSE}
ttops <- find_trees(point_cloud_norm, lmf( ws = 5))
```

```{r p16, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
plot(chm, col = height.colors(50))
plot(ttops, add = TRUE)
```

```{r p17, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
x <- plot(point_cloud_norm, bg = "white", size = 4, axis = TRUE, legend = TRUE)
add_treetops3d(x, ttops)
```

## Individual Tree Segmentation
```{r p18, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
algo <- dalponte2016(chm, ttops)
point_cloud_norm <- segment_trees(point_cloud_norm, algo) 
plot(point_cloud_norm, bg = "white", size = 3, color = "treeID") 
```

```{r p19, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE, rgl=TRUE}
plot(filter_poi(point_cloud_norm, treeID == 5), size = 3, bg = "white",  axis = TRUE, legend = TRUE)
```

## Individual Tree Metrics
```{r p20, echo=TRUE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
ttops <- tree_metrics(point_cloud_norm, func = .stdmetrics) %>% st_as_sf()
kable(summary(ttops[,c(	"zmax",	"zmean","zsd","zq95","area")]%>% st_drop_geometry()))
hist(ttops$zmax,plot = T, main = "Height Distribruition (m)",xlab =   "Height (m)",ylab = "Frequency")
```


